---
title: "formal analysis - random forest ALE, SCYM data"
author: "jill deines"
date: "03/20/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**Goal: use random forests to identify effects of WT using the SCYM yield data**


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, fig.path='../figure/03.03_analysis_main_rf_ale/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE, eval=TRUE}
library(dplyr)
library(ggplot2)
library(readr)
library(earth)
library(randomForest)

library(here)
library(patchwork)



sessionInfo()
```

# Directories

```{r directories, message = FALSE}
# for master file
masterDataDir <- '/Users/dein121/Library/CloudStorage/OneDrive-PNNL/Documents/projects/2022_compass_GLM/paper_gw_subsidy/data/RepoData'

# intermediate folder for large objects (eg rf output)
scratchFolder <- paste0(masterDataDir,'/intermediates')
# make scratch folder if necessary
dir.create(file.path(scratchFolder), showWarnings = FALSE)

# output folder: clean df for figures
repoDataDir <- paste0(here::here(),'/data/analysisOutput_forFigs')

```

# functions
helper functions for processing ALEplot output

```{r funs_ale}
library(ALEPlot)
aleToDf <- function(aleList, xname){
  df <- data.frame(xname = aleList$x.values,
                      'YieldEffect' = aleList$f.values)
  names(df)[1] <- xname
  return(df)
}

# prep prediction function for ALEplot
yhat <- function(X.model, newdata) as.numeric(predict(X.model, newdata, type = 'response'))

```

# SCYM

## Variable selection

load master data file and use MARS for variable selection

note water deficit is named "spei" in the df

RUN 1X - OUTPUT STORED AS RDS FOR LATER USE

```{r variableSelection, eval = FALSE}
master <- readRDS(paste0(masterDataDir,'/masterData_pointSample.rds'))

# identify variables to consider (listed in Table S1)

df_table1var <- master %>%
  dplyr::select(c(yield_tha, nccpi3corn, rootznaws, JJAradn, year, pr_early, pr_grow,
                  GDD_8_30, spei_ju1, spei_au2, soc0100, drclasscd, contains('tmin'),
                   contains('tmax'), contains('vpd'), contains('ppt'), tpi,  slope,
                   psim_07_m, psim_05_m, psim_06_m, psim_08_m, psim_14_m)) %>%
  tidyr::drop_na()

# mars for var selection
set.seed(5)
marsAll <- earth::earth(yield_tha ~ ., data = df_table1var, degree = 2)
summary(marsAll)
ev <- earth::evimp(marsAll) # estimate variable importance
plot(ev)
ev

saveRDS(marsAll, paste0(scratchFolder, '/marsOutput_scymDf_forVarSel.rds'))
```



## RF model
run 1x - output stored as RDS for later use

Note: July soil moisture from the Standard Crop Model is added as a variable based on helpful reviewer comments and strategy to examine direct vs soil moisture pathways for groundwater contributions

```{r scym_rf, eval = FALSE}
# load MARS output for selected variables
marsAll <- readRDS( paste0(scratchFolder,'/marsOutput_scymDf_forVarSel.rds'))
summary(marsAll)
ev <- earth::evimp(marsAll) # estimate variable importance
plot(ev)
ev

# vector of variable names
envvars0 <- rownames(ev)[1:17]


# load master df and add in july soil moisture from psim without wt
scym_df0 <- readRDS(paste0(masterDataDir,'/masterData_pointSample.rds'))


# get apsim soil moisture: standard model
apsim_sm <- read_csv(paste0(masterDataDir,'/standardModel_JulySoilMoisture.csv'))


scym_df <- scym_df0 %>%
  left_join(apsim_sm)

# add standard model soil moistures to variable vector
envvars <- c(envvars0, 'SW1m_Jul')


# prep random forest - df with only wanted variables
df_20201  <- scym_df %>%
  dplyr::select(c(yield_tha, all_of(envvars), spei_ju1))

# run and save RF
set.seed(5)
ncols <- dim(df_20201)[2]
rf_scym20 <- randomForest(x = df_20201[,2:ncols],
                         y = df_20201$yield_tha,
                         ntree = 100,
                         importance = FALSE,
                         nodesize = 10)
rf_scym20
saveRDS(rf_scym20, paste0(scratchFolder,'/rf1_scymPointRf_marsVars_standardSoilMoisture_ns10_100t.RDS'))


```


## scym  aleplots - 

```{r aleTest_scym2020}
# load data and intermediate model outputs --------------
marsAll <- readRDS( paste0(scratchFolder,'/marsOutput_scymDf_forVarSel.rds'))
ev <- earth::evimp(marsAll) # estimate variable importance
envvars0 <- rownames(ev)[1:17]
envvars <- c(envvars0, 'SW1m_Jul', 'spei_ju1')

# df: master plus Standard Model Soil Moisture
scym_df0 <- readRDS(paste0(masterDataDir,'/masterData_pointSample.rds')) %>% ungroup()
apsim_sm <- read_csv(paste0(masterDataDir,'/standardModel_JulySoilMoisture.csv'))
scym_df <- scym_df0 %>%
  left_join(apsim_sm)

mean(scym_df$yield_tha)

# df
Xdat_scym  <- scym_df %>%
  dplyr::select(c(yield_tha, all_of(envvars))) %>%
  tidyr::drop_na() %>%
  as.data.frame()

# rf model
rfModel_scym <- readRDS(paste0(scratchFolder,'/rf1_scymPointRf_marsVars_standardSoilMoisture_ns10_100t.RDS'))

rfModel_scym



# ALE plot for WT variable (psim_07_m, column = J = 17)
names(Xdat_scym)

ALEPlot_scym <- ALEPlot(X = Xdat_scym, X.model = rfModel_scym, J = 17, 
                        pred.fun = yhat, K = 100)

# rearrange and visualize
scymYield_df <- aleToDf(ALEPlot_scym, 'psim7') 

ggplot(scymYield_df, aes(x = psim7, y = YieldEffect)) +
  geom_line() + geom_rug() + theme_bw()
max(scymYield_df$YieldEffect)



# save data for final fig
write_csv(scymYield_df, paste0(repoDataDir,'/Fig3a_scym_yieldVsWT_aleplot_may2024_35kNoDupe_standardSoilM.csv'))

```

## BOOTSTRAP
figure S1b

```{r aleBOOT_scym, eval = FALSE}
nsamples <- 200
n_boot <- nrow(Xdat_scym)
boots <- list()
for (i in 1:nsamples){
  # resample
  resampled <- Xdat_scym %>%
    slice_sample(n = n_boot, replace = TRUE)
  
  ale <- ALEPlot(X = resampled, X.model = rfModel_scym,
                            J = 17, pred.fun = yhat, K = 100)
  ale_df <- aleToDf(ale, 'psim7') 
  
  df_out <- ale_df %>% mutate(bootN = i)
  
  boots[[i]] <- df_out
}



allBoots <- do.call('rbind', boots)

ggplot(allBoots,
       aes(x = psim7, y = YieldEffect, group = as.factor(bootN), color =  bootN)) +
  geom_line() + theme_bw() + scale_color_viridis_c()


# save data for final fig
write_csv(allBoots, paste0(repoDataDir,'/Fig3a_Bootstrap_',nsamples,
                           '_may2024_standardSoilM.csv'))

```

### analyze bootstrap
```{r bootstrap, dpi = 300, fig.width = 3.5, fig.height = 3}
nsamples <- 200
allBoots <- read_csv(paste0(repoDataDir,'/Fig3a_Bootstrap_',nsamples,
                           '_may2024_standardSoilM.csv'))


p_bootStrap <- ggplot(allBoots,
       aes(x = psim7, y = YieldEffect, group = as.factor(bootN))) +
  geom_line(alpha = .15) + 
  geom_hline(yintercept = -0.025, linetype = 'dashed', color = 'red') +
    geom_vline(xintercept = 1.47, linetype = 'dashed', color = 'blue') +
   # geom_hline(yintercept = .05) +
      geom_hline(yintercept = 0, linetype = 'dotted') +
  theme_bw() + xlab('Depth to Water Table (m)') + ylab('Yield Effect (t/ha)') +
  scale_color_viridis_c()
p_bootStrap

# get maxima
boot_max <- allBoots %>%
  group_by(bootN) %>%
  slice_max(YieldEffect) %>%
  mutate(effectSize = YieldEffect + 0.025)
quantile(boot_max$effectSize, c(.025, .975)) 
```


## Figure 3d: wt by spei scym

```{r scym_wtByClim_agu}

#  by climate ---------------------
Xdat3 <- Xdat_scym %>%
  mutate(spei_tile = ntile(spei_ju1, 5),
         vpd_tile = ntile(vpd_july, 5),
        speiJu_class = case_when(spei_ju1 >= 0 ~ 1,
                                spei_ju1 < 0  & spei_ju1 >= -50 ~ 2,
                                spei_ju1 < -50 & spei_ju1 >= -150 ~ 3,
                                spei_ju1 < -150 ~ 4)) %>%
  mutate(speiClass = case_when(speiJu_class == 1 ~ 'wet',
                               speiJu_class == 2 ~ 'normal',
                               speiJu_class == 3 ~ 'dry',
                               speiJu_class == 4 ~ 'very dry'),
         gwClass = case_when(psim_07_m < 1.1 ~ 1,
                             psim_07_m >= 1.1 & psim_07_m <2.5 ~ 2,
                             psim_07_m >=2.5 ~ 3),
         yearGroup = case_when(year < 2009 ~ 1,
                               year >= 2009 ~ 2))

# mean yield in dry
table(Xdat3$speiJu_class)
veryDry <- Xdat3 %>% filter(speiJu_class == 4)
summary(veryDry$yield_tha)


names(Xdat3)
psim_speiList <- list()
for(i in 1:4) {
  y <- i
  df_year <- Xdat3 %>% filter(speiJu_class == y)
  ale_year <- aleToDf(ALEPlot(X = df_year, X.model = rfModel_scym, J = 17, pred.fun = yhat, K = 40), 'psim_07_m')
  ale_year$speiJu_class <- y
  psim_speiList[[i]] <- ale_year
}
psim_spei <- do.call('rbind', psim_speiList)



# save data for final fig
write_csv(psim_spei, paste0(repoDataDir,'/Fig3d_scym_yieldVsWT_byPPET_aleplot_oct23_35knoDupe.csv'))


aguP <- ggplot(psim_spei,
       aes(x = psim_07_m, y = YieldEffect, group = as.factor(speiJu_class), 
           color =  as.factor(speiJu_class))) +
  geom_line() + 
    geom_hline(yintercept = 0, linetype = 'dashed') +
  scale_color_manual(values = c('blue','gray40','#f4a582','red'),
                     labels = c('wet','normal','drier','very dry')) +
  ylab('Yield Effect (t/ha)') + xlab('July Depth to Water (m)') +
  theme_bw() +  theme(legend.title = element_blank(),
                      legend.position = c(.7,.3),
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank(),
                     text = element_text(size = 14)) 
aguP

```



# psim WT model

## random forest

```{r psim_rf, eval = FALSE}
psimDiffs0 <- readRDS(paste0(cleanDir,'/intermediateGridPoints/hydroCorn_masterDataGrid_v02_V05_scym9_detrended_withPsims_weatherGroups_LAI_19992020.rds'))

# same eyar as scym
psim <- psimDiffs0 %>% filter(year < 2019 & scym2020_10prct == 1) %>%
  dplyr::select(-contains('SW1')) %>%
  left_join(psim_sm %>% dplyr::select(psim_id, year, SW1m_Jul))

# load mars
marsAll <- readRDS( paste0(scratchFolder0,'/scymP_mars_c1_35kSub_train2_scym2020.rds'))
ev <- earth::evimp(marsAll) # estimate variable importance
varsWanted <- rownames(ev)[1:17]
varsWanted2 <- varsWanted[!grepl('psim_07_m',varsWanted)]

varsWanted3 <- c(varsWanted2, 'WT_Jul_m','spei_ju1','SW1m_Jul')

# prep df for rf
psim_df  <- psim %>%
  dplyr::select(c(psim_WT_tha, all_of(varsWanted3) )) %>% tidyr::drop_na()
nrow(psim_df)

set.seed(5)
ncols <- dim(psim_df)[2]
rf_psim <- randomForest(x = psim_df[,2:ncols],
                         y = psim_df$psim_WT_tha,
                         ntree = 100,
                         importance = FALSE,
                        nodesize = 10)
rf_psim
saveRDS(rf_psim, paste0(scratchFolder,'/rf1_psim_marsVars_spei_WT_ns10_t100_May2024_standardSoilM.RDS'))
```

## psim ale plots

```{r psimAle}
psimDiffs0 <- readRDS(paste0(cleanDir,'/intermediateGridPoints/hydroCorn_masterDataGrid_v02_V05_scym9_detrended_withPsims_weatherGroups_LAI_19992020.rds'))


# load mars
marsAll <- readRDS( paste0(scratchFolder,'/scymP_mars_c1_35kSub_train2_scym2020.rds'))
ev <- earth::evimp(marsAll) # estimate variable importance
varsWanted <- rownames(ev)[1:17]
varsWanted2 <- varsWanted[!grepl('psim_07_m',varsWanted)]

varsWanted3 <- c(varsWanted2, 'WT_Jul_m','spei_ju1')


# same eyar as scym
Xdat <- psimDiffs0 %>%
  filter(year < 2019 & scym2020_10prct == 1) %>%
  dplyr::select(-contains('SW1')) %>%
  left_join(psim_sm %>% dplyr::select(psim_id, year, SW1m_Jul)) %>%
  dplyr::select(c(psim_WT_tha, psim_noWT_tha, all_of(varsWanted3) )) %>%
  mutate(sim_diff = psim_WT_tha - psim_noWT_tha) %>%
  tidyr::drop_na() %>%
  as.data.frame()

# model
rfModel <- readRDS( paste0(scratchFolder,'/rf1_psim_marsVars_spei_WT_ns10_t100_May2024_standardSoilM.RDS'))
rfModel

names(Xdat)

# psim - change with WT
ALEPlot_psim <- ALEPlot(X = Xdat, X.model = rfModel, J = 19, pred.fun = yhat, K = 100)

psimYield_df <- aleToDf(ALEPlot_psim, 'psim7')

ggplot(psimYield_df, aes(x = psim7, y = YieldEffect)) +
  geom_line() + geom_rug() + theme_bw()

psimPlot <- ggplot(psimYield_df,
       aes(x = psim7, y = YieldEffect)) +
  # geom_rect(xmin = .775, xmax = 2.48, ymin = -Inf, ymax = Inf, fill = 'lightblue') +
  geom_line() +
  geom_rug() +
  coord_cartesian(ylim=c(-.325, 0.4)) +
  geom_hline(yintercept = 0, linetype = 'dotted') +
  geom_hline(yintercept = -.035, linetype = 'dashed', color = 'red') +
  # geom_vline(xintercept = 1.23, linetype = 'dashed', color = 'blue') +
    xlab('July Depth to Water (m)') + ylab('Yield Effect (t/ha)') +
  theme_bw()+ theme(legend.position = 'none',
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank()) +
  ggtitle('PSIMS Simulations')
psimPlot

# simulated difference
ggplot(Xdat, aes(x = WT_Jul_m, y = sim_diff)) +
  geom_smooth()


# # change with precip
# ALEPlot_pr <- ALEPlot(X = Xdat, X.model = rfModel, J = 8, pred.fun = yhat, K = 100)
# 
# psimYield_df_pr <- aleToDf(ALEPlot_pr, 'pr_grow') %>%
#   mutate(sd = sd(psimDiffs0 %>% 
#                    filter(year < 2019 & scym2020_10prct == 1) %>%
#                    pull(pr_grow))) %>%
#   mutate(Effect_sd = YieldEffect / sd)
# 
# ggplot(psimYield_df_pr, aes(x = pr_grow, y = YieldEffect)) +
#   geom_line() + geom_rug() + theme_bw()
# 
# # change with july vpd
# ALEPlot_vpd <- ALEPlot(X = Xdat, X.model = rfModel, J = 3, pred.fun = yhat, K = 100)
# 
# psimYield_df_vpd <- aleToDf(ALEPlot_vpd, 'vpd_july') 
# 
# ggplot(psimYield_df_vpd, aes(x = vpd_july, y = YieldEffect)) +
#   geom_line() + geom_rug() + theme_bw()
# 
# 
# # change with nccpi
# ALEPlot_nccpi <- ALEPlot(X = Xdat, X.model = rfModel, J = 7, pred.fun = yhat, K = 100)
# 
# psimYield_df_nccpi <- aleToDf(ALEPlot_nccpi, 'nccpi') 
# 
# ggplot(psimYield_df_nccpi, aes(x = nccpi, y = YieldEffect)) +
#   geom_line() + geom_rug() + theme_bw()

write_csv(psimYield_df,paste0(repoDataDir,'/Fig3b_psim_yieldVsWT_aleplot_may2024_standardSoilM.csv'))

```

# psim WT model - scym yields

## random forest

```{r psim_rf_scym, eval = FALSE}
psimDiffs0 <- readRDS(paste0(cleanDir,'/intermediateGridPoints/hydroCorn_masterDataGrid_v02_V05_scym9_detrended_withPsims_weatherGroups_LAI_19992020.rds'))

# get psim soil moisture
cleanDir2 <- '/Users/dein121/Library/CloudStorage/OneDrive-PNNL/Documents/projects/2022_compass_GLM/paper_gw_subsidy/data/psims/processed'
psim_noWT <- readRDS(paste0(cleanDir2,'/psims_noWT_1984-2020_20240518.rds'))
psim_sm <- psim_noWT %>% dplyr::select(psim_id,  year, contains('SW1'))

# same eyar as scym
psim <- psimDiffs0 %>% filter(year < 2019 & scym2020_10prct == 1) %>%
  dplyr::select(-contains('SW1')) %>%
  left_join(psim_sm %>% dplyr::select(psim_id, year, SW1m_Jul))

# load mars
marsAll <- readRDS( paste0(scratchFolder0,'/scymP_mars_c1_35kSub_train2_scym2020.rds'))
ev <- earth::evimp(marsAll) # estimate variable importance
varsWanted <- rownames(ev)[1:17]
varsWanted2 <- varsWanted[!grepl('psim_07_m',varsWanted)]

varsWanted3 <- c(varsWanted2, 'WT_Jul_m','spei_ju1','SW1m_Jul')

# prep df for rf
psim_df  <- psim %>%
  dplyr::select(c(yield_tha, all_of(varsWanted3) )) %>% tidyr::drop_na()
nrow(psim_df)

set.seed(5)
ncols <- dim(psim_df)[2]
rf_psim <- randomForest(x = psim_df[,2:ncols],
                         y = psim_df$yield_tha,
                         ntree = 100,
                         importance = FALSE,
                        nodesize = 10)
rf_psim
saveRDS(rf_psim, paste0(scratchFolder,'/rf1_psim_marsVars_spei_WT_ns10_t100_May2024_standardSoilM_scymVersion.RDS'))
```

## psim ale plots - scymv ersion

```{r psimAle_scym}
psimDiffs0 <- readRDS(paste0(cleanDir,'/intermediateGridPoints/hydroCorn_masterDataGrid_v02_V05_scym9_detrended_withPsims_weatherGroups_LAI_19992020.rds'))


# load mars
marsAll <- readRDS( paste0(scratchFolder0,'/scymP_mars_c1_35kSub_train2_scym2020.rds'))
ev <- earth::evimp(marsAll) # estimate variable importance
varsWanted <- rownames(ev)[1:17]
varsWanted2 <- varsWanted[!grepl('psim_07_m',varsWanted)]

varsWanted3 <- c(varsWanted2, 'WT_Jul_m','spei_ju1','SW1m_Jul')

# get psim soil moisture
cleanDir2 <- '/Users/dein121/Library/CloudStorage/OneDrive-PNNL/Documents/projects/2022_compass_GLM/paper_gw_subsidy/data/psims/processed'
psim_noWT <- readRDS(paste0(cleanDir2,'/psims_noWT_1984-2020_20240518.rds'))
psim_sm <- psim_noWT %>% dplyr::select(psim_id,  year, contains('SW1'))


# same eyar as scym
Xdat <- psimDiffs0 %>%
  filter(year < 2019 & scym2020_10prct == 1) %>%
  dplyr::select(-contains('SW1')) %>%
  left_join(psim_sm %>% dplyr::select(psim_id, year, SW1m_Jul)) %>%
  dplyr::select(c(yield_tha, all_of(varsWanted3) )) %>%
  tidyr::drop_na() %>%
  as.data.frame()

# model
rfModel <- readRDS( paste0(scratchFolder,'/rf1_psim_marsVars_spei_WT_ns10_t100_May2024_standardSoilM_scymVersion.RDS'))
rfModel

names(Xdat)

# psim - change with WT
ALEPlot_psim <- ALEPlot(X = Xdat, X.model = rfModel, J = 18, pred.fun = yhat, K = 100)

psimYield_df <- aleToDf(ALEPlot_psim, 'psim7')

ggplot(psimYield_df, aes(x = psim7, y = YieldEffect)) +
  geom_line() + geom_rug() + theme_bw()

scym <- ggplot(psimYield_df,
       aes(x = psim7, y = YieldEffect)) +
   geom_rect(xmin = 1.05, xmax = 2.52, ymin = -Inf, ymax = Inf, fill = 'lightblue') +
  geom_line() +
  geom_rug() +
  coord_cartesian(ylim=c(-.325, 0.4)) +
  geom_hline(yintercept = 0, linetype = 'dotted') +
  geom_hline(yintercept = -.025, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = 1.47, linetype = 'dashed', color = 'blue') +
    xlab('July Depth to Water (m)') + ylab('Yield Effect (t/ha)') +
  theme_bw()+ theme(legend.position = 'none',
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank()) +
  ggtitle('SCYM Observations')

scym

max(psimYield_df$YieldEffect)
psimYield_df %>%
  slice_max(YieldEffect)
max(psimYield_df$YieldEffect) + 0.025



# simulated difference
ggplot(Xdat, aes(x = WT_Jul_m, y = sim_diff)) +
  geom_smooth()


# # change with precip
# ALEPlot_pr <- ALEPlot(X = Xdat, X.model = rfModel, J = 8, pred.fun = yhat, K = 100)
# 
# psimYield_df_pr <- aleToDf(ALEPlot_pr, 'pr_grow') %>%
#   mutate(sd = sd(psimDiffs0 %>% 
#                    filter(year < 2019 & scym2020_10prct == 1) %>%
#                    pull(pr_grow))) %>%
#   mutate(Effect_sd = YieldEffect / sd)
# 
# ggplot(psimYield_df_pr, aes(x = pr_grow, y = YieldEffect)) +
#   geom_line() + geom_rug() + theme_bw()
# 
# # change with july vpd
# ALEPlot_vpd <- ALEPlot(X = Xdat, X.model = rfModel, J = 3, pred.fun = yhat, K = 100)
# 
# psimYield_df_vpd <- aleToDf(ALEPlot_vpd, 'vpd_july') 
# 
# ggplot(psimYield_df_vpd, aes(x = vpd_july, y = YieldEffect)) +
#   geom_line() + geom_rug() + theme_bw()
# 
# 
# # change with nccpi
# ALEPlot_nccpi <- ALEPlot(X = Xdat, X.model = rfModel, J = 7, pred.fun = yhat, K = 100)
# 
# psimYield_df_nccpi <- aleToDf(ALEPlot_nccpi, 'nccpi') 
# 
# ggplot(psimYield_df_nccpi, aes(x = nccpi, y = YieldEffect)) +
#   geom_line() + geom_rug() + theme_bw()

write_csv(psimYield_df,paste0(repoDataDir,'/Fig3b_psim_yieldVsWT_aleplot_may2024_standardSoilM.csv'))

```

# All Together

AGU ALE plot

## scym only



```{r Fig_scymSubsidy, fig.width = 3.25, fig.height = 4, dpi = 300, dev = c('pdf','png')}

sd_master <- scymYield_df %>% mutate(model = '1SCYM point') %>%
  bind_rows(psimYield_df %>% mutate(model = '2PSIM Yield'))

ggplot(sd_master %>% filter(model == '1SCYM point'),
       aes(x = psim7, y = YieldEffect)) +
   geom_rect(xmin = 1.05, xmax = 2.55, ymin = -Inf, ymax = Inf, fill = 'lightblue') +
  geom_line() +
  geom_rug() +
  geom_hline(yintercept = 0, linetype = 'dotted') +
  geom_hline(yintercept = -.035, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = 1.47, linetype = 'dashed', color = 'blue') +
    xlab('July Depth To Water (m)') + ylab('Yield Effect (t/ha)') +
  theme_bw()+ theme(legend.position = 'none',
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank())


```

## psim only

```{r Fig_psimSubsidy, fig.width = 3.25, fig.height = 4, dpi = 300, dev = c('pdf','png')}

sd_master <- scymYield_df %>% mutate(model = '1SCYM point') %>%
  bind_rows(psimYield_df %>% mutate(model = '2PSIM Yield'))

ggplot(sd_master %>% filter(model == '2PSIM Yield'),
       aes(x = psim7, y = YieldEffect)) +
   geom_rect(xmin = .75, xmax = 2.55, ymin = -Inf, ymax = Inf, fill = 'lightblue') +
  geom_line() +
  geom_rug() +
  coord_cartesian(ylim = c(-.75,.6)) +
  geom_hline(yintercept = 0, linetype = 'dotted') +
  geom_hline(yintercept = -.07, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = 1.25, linetype = 'dashed', color = 'blue') +
    xlab('July Depth To Water (m)') + ylab('Yield Effect (t/ha)') +
  theme_bw()+ theme(legend.position = 'none',
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank())


```

## scym and psim

```{r agu_psim_scym, fig.height = 5, fig.width = 3.25, dpi = 300, dev = c('pdf','png')}
sd_master <- scymYield_df %>% mutate(model = '1SCYM point') %>%
  bind_rows(psimYield_df %>% mutate(model = '2PSIM Yield'))

aguPlot <- ggplot(sd_master,
       aes(x = psim7, y = YieldEffect)) +
  geom_line() +
  facet_wrap(~model, nrow = 2, scales = 'free_y') +
  xlab('July Depth To Water (m)') + ylab('Yield Effect (t/ha)') +
  geom_hline(yintercept = 0, linetype = 'dotted') +
  theme_bw()+ theme(legend.position = 'none',
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank(),
                    strip.background = element_blank(),
  strip.text.x = element_blank(),
                     text = element_text(size = 14)) 
aguPlot

# ggsave(filename = '/Users/deinesji/Dropbox/4Deines/miscPresentations/20211210_AGU21/talk_ExtremeEvents/input/aleplot.png',
#        plot = aguPlot,
#        width = 3.25, height = 5, units = 'in')
```


Fig manuscript

```{r Fig_scymPsimSubsidy, fig.width = 6.5, fig.height = 4, dpi = 300, dev = c('pdf','png')}

ggplot(sd_master ,#%>% filter(model == '1SCYM point'),
       aes(x = psim7, y = YieldEffect)) +
   geom_rect(xmin = 1.05, xmax = 2.75, ymin = -Inf, ymax = Inf, fill = 'lightblue') +
  geom_line() +
  geom_rug() +
  geom_hline(yintercept = 0, linetype = 'dotted') +
  geom_hline(yintercept = -.0425, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = 1.5, linetype = 'dashed', color = 'blue') +
    xlab('July Depth To Water (m)') + ylab('Yield Effect (t/ha)') +
  facet_wrap(~model, scales = 'free_y') +
  theme_bw()+ theme(legend.position = 'none',
                      panel.grid.major.x = element_blank(),
                      panel.grid.major.y = element_blank(),
                     panel.grid.minor.y = element_blank(),
                      panel.grid.minor.x = element_blank())


```







